---
title: "Chapter 4: Design & Analysis of Factorial Experiments"
author: "Dr. Katie Fitzgerald"
output: 
    xaringan::moon_reader:
      css: [default-fonts, default, boxes.css]
      yolo: false
      includes:
        after_body: insert-logo.html  
      seal: false  
      lib_dir: libs
      nature:
        highlightStyle: github
        highlightLines: true
        countIncrementalSlides: false
        ratio: "16:9"
---

```{r child = "setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
```

class: middle, center

# Chapter 4: Design & Analysis of Factorial Experiments

---

# AE-CH4

+ Navigate to course Github page
+ Open AE-CH4.qmd and copy contents
+ Create new .qmd in your MATH_350 project folder and paste contents from GitHub
+ Render your .qmd

---

# The process of statistics

+ Formulating well-defined questions
+ Properly designing a study
+ Collecting meaningful data
+ Deciding how to properly analyze and draw conclusions from data

---

# Case study: which popcorn brand is best?

Want to test the quality to two microwave popcorn brands. Could test with a two-sample t-test, but would have to assume that many other factors are held constant. What other factors (besides brand) might affect the quality of popcorn? 

--

In **Factorial designs** are structured to use factors (i.e. explanatory variables) to organize meaningful groups of treatment conditions and simultaneously test the effects of multiple factor-level combinations on a response variable of interest.

--

Two-sample t-test is a special case of a factorial design that has just one factor and two levels (e.g. Brand, levels A and B).

---

# Elements of a Well-Designed Experiment

+ Attempts to answer the question(s) of interest as clearly and efficiently as possible. Any statistical analysis is only as good as the quality of the data.

    + What measurements will be taken?
    + What material is needed? 
    + What procedures will be used?
    + What background knowledge is needed?

1. Clearly define the problem and objective of the experiment
2. Identify a response variable
3. Identify the factors and levels
4. Choose a design
5. Determine a sample size

---

# Identifying a Response Variable

What are possible ways of measuring "quality" of the popcorn?

--

+ taste rating
+ texture rating
+ volume of the kernels
+ Number of popped kernels
+ Number of unpopped kernels remaining

--

They chose % of kernels that were popped. Why is this important?

--

Considered only looking at a random sample of 20 kernels, but decided against it. Was this a good move? 

---

# Identifying the Factors and Levels

+ Factor 1: popcorn `Brand` at two levels, Fastco (relatively cheap) and Pop Secret (relatively expensive)
+ Factor 2: `Microwave` at two levels, dorm lounge and dorm room
+ Factor 3: cooking `Time` at two levels, 105 seconds and 135 seconds

Held Constant:

+ Highest power setting on both microwaves
+ Scored all popcorn on a shelf in their room
+ Butter flavor of each brand

Could not control:

+ age of popcorn
+ manufacturing plant
+ storage before purchase

Attempt to take random sample of bags, and to randomly allocate bags to factor-level conditions, to control for uncontrollable factors. 

---

# Choosing a Design

**Experimental units** are the smallest part of the experimental material that is randomly assigned to a factor-level condition. Here, the units are the bags of popcorn. 

--

If each unit is randomly assigned to just one condition, and all factor-levels are **crossed** (each level of a factor is tested in combination with each level of the other factors), this is called a **completely randomized factorial design**. 

--

Block designs, split-plot (repeated measures), and nested designs are more complex alternatives (Chapter 5).

---

# Sample size

More power to detect differences between groups with larger sample. But larger sample costs more (time and $). 

+ Estimate they can count one bag in 10-15 minutes
+ Can recruit friends to count too (concern?)

--

Can conduct 32 tests in about 4 hours. Why not 30? 

--

**Balanced design** has same number of tests/trials for each factor-level combination. This will result in a more powerful test. 

---

## Analyzing Two-Way ANOVA

|Source | DF | SumSq | MeanSq | F | p-value
|-------|-------|-------|-------|-------|-------|
|Factor A | (a - 1) | SSA | MSA | MSA/MSE
|Factor B | (b - 1) | SSB | MSB | MSB/MSE
|AxB      | (a -1)(b -1)| SSAB | MSAB | MSAB / MSE
|Residuals | (n - ab) | SSE | MSE

Variability in Y can be broken down into 
+ Variability between groups in Factor A (SSA)
+ Variability between groups in Factor B (SSB)
+ Variability between groups in FactorAxB combinations (SSAB)
+ Variability within groups that's left unexplained by A, B, and AB.

+ The F-statistics are the ratios of the **between-group** variability and the **within-group** variability
+ When $H_0$ is true (no differences between groups), the F ratio will be close to 1
+ As the differences between groups increases ($H_A$ true), the F-ratio grows large
+ We reject $H_0$ for large F values

---

# Analyzing Two-Way ANOVA

$$y_{ijk} = \mu + \alpha_i + \beta_j + (\alpha\beta)_{ij} + \epsilon_{ijjk}$$

$$i = 1,2; j = 1,2; k = 1,2,\dots 8$$

There are three sets of hypotheses

$$H_{01}: \mu_{\text{Fastco}} = \mu_{\text{PopSecret}}$$
$$H_{A1}: \mu_{\text{Fastco}} \neq \mu_{\text{PopSecret}}$$

$$H_{02}: \mu_{105} = \mu_{135}$$
$$H_{A2}: \mu_{103} \neq \mu_{135}$$

$$H_{03}: \text{There is no interaction between Time and Brand}$$
$$H_{A3}: \text{There IS an interaction between Time and Brand}$$

Saying there is no interaction $(H_{03})$ is equivalent to saying that `Time` has no influence on how `Brand` affects `pop_rate`. Or equivalently, `Brand` has no influence on how `Time` affects `pop_rate`. 

