---
title: "Exam 1 Review"
subtitle: "Chapters 1,2,4"
author: "Dr. Katie Fitzgerald"
output: 
    xaringan::moon_reader:
      css: [default-fonts, default, boxes.css]
      yolo: false
      includes:
        after_body: insert-logo.html  
      seal: false  
      lib_dir: libs
      nature:
        highlightStyle: github
        highlightLines: true
        countIncrementalSlides: false
        ratio: "16:9"
---

```{r child = "setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
library(janitor)
```

class: middle, center

# Exam 1 Review: Chapters 1, 2, 4

---

# Homework revisited

**Chapter 2, E.2** Suppose a test for $H_0: \beta_1 = 0$ vs. $H_A: \beta_1 \neq 0$ returns a small p-value. What CAN we conclude, and what can we NOT conclude? Why? If the answer is nuanced or "it depends", indicate what information you would need to draw a more clear conclusion. 

  a) Reject $H_0$
  
  b) $\mu_1 - \mu_2 \neq 0$
  
  c) model provides a good estimate of response for a given value of $x_i$
  
  d) $x$ is a significant predictor of $y$
  
  e) $x$ causes $y$ to increase
  
  f) there is a linear relationship between $x$ and $y$
  

---

# Homework revisited

**Chapter 2, E.4** How would we 