---
title: "HW Bayes 03"
author: "Your Name Here"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number_section: false
    highlight: "tango"
    theme: "cosmo"
    df-print: paged
editor: visual
editor_options: 
  chunk_output_type: console
---

## Instructions

This Homework comes from Chapters 3 & 8 of the [Bayes Rules! Textbook](https://www.bayesrulesbook.com/).

### [Chapter 3 Exercises](https://www.bayesrulesbook.com/chapter-3#exercises):

-   3.1

-   3.3

-   3.4

-   3.11

-   3.18

-   BONUS: 3.8

### [Chapter 8 Exercises:](https://www.bayesrulesbook.com/chapter-8#exercises-1)

-   8.2

-   8.4

-   8.6

-   8.9

-   8.14 (skip a, use a beta prior)

```{r}
#| label: load-packages
#| message: false
#| warning: false

```

## 3.1

#### Part a

> YOUR ANSWER HERE

#### Part b

> YOUR ANSWER HERE

#### Part c

> YOUR ANSWER HERE

#### Part d

> YOUR ANSWER HERE

## 3.3

#### Part a

> YOUR ANSWER HERE

```{r}

```


#### Part b

> YOUR ANSWER HERE

#### Part c

> YOUR ANSWER HERE

#### Part d

> YOUR ANSWER HERE

```{r}

```


#### Part e

> YOUR ANSWER HERE

```{r}

```


## 3.4

> YOUR ANSWER HERE

## 3.11

### Part a

> YOUR ANSWER HERE

```{r}

```

### Part b

> YOUR ANSWER HERE

### Part c

```{r}

```

### Part d

```{r}

```

> YOUR ANSWER HERE

## 3.18

### Part a

```{r}

```

### Part b

```{r}

```


### Part c

> YOUR ANSWER HERE

## 3.8 (BONUS)

*Can upload screenshot of by-hand work, or type up derivations in LaTeX*

## 8.2

### Part a

> YOUR ANSWER HERE

### Part b

> YOUR ANSWER HERE

## 8.4 

### Part a

> YOUR ANSWER HERE

### Part b

> YOUR ANSWER HERE

### Part c

> YOUR ANSWER HERE

## 8.6

### Part a

```{r}

```

### Part b

```{r}

```

### Part c

```{r}

```

## 8.9

### Part a

```{r}

```

### Part b

```{r}

```

> YOUR ANSWER HERE

### Part c

```{r}

```

> YOUR ANSWER HERE

### Part d

```{r}

```

> YOUR ANSWER HERE

### Part e

> YOUR ANSWER HERE

## 8.14

### Part b

> YOUR ANSWER HERE

### Part c

> YOUR ANSWER HERE

### Part d

```{r}

```

### Part e

```{r}

```

> YOUR ANSWER HERE

## Reflection prompts

*Respond to each of the following after you have completed all exercises in this assignment*

*(RP1): What were the main concepts covered in this assignment?*

> YOUR ANSWER HERE

*(RP2): What's one thing you understand better after completing these exercises?*

> YOUR ANSWER HERE

*(RP3): What exercise(s) gave you the most trouble? What was difficult about them/where did you get stuck?*

> YOUR ANSWER HERE
