---
title: "AE-CH8 / HW-CH8"
author: "Your Name Here"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number_section: false
    highlight: "tango"
    theme: "cosmo"
    df-print: paged
editor: visual
execute: 
  warning: false
  message: false
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE}
options(scipen=999)
```


This Application Exercise walks through the guided activities in Chapter 8 of the textbook. There are spaces for you to answer each question (with code and/or text). Feel free to also use this document to add your own notes along the way.

## Load packages & data

```{r}
#| label: load-packages-data
#| message: false
#| warning: false
library(tidyverse)
library(janitor)
library(broom)
theme_set(theme_minimal())

cancer <- read_csv("./data/C8 CancerCluster.csv") %>% 
  clean_names()
```

# Cancer data

## Question 1

```{r}
67/1138
```

Proportion of people who reported cancer diagnosis was 67/1138 = `r round(67/1138, 4)`

## Question 2

```{r}
person_years <- 1138*25
```

Number of person-years was `r person_years`

## Question 3

```{r}
cancer_rate <- 67/person_years
cir <- cancer_rate*100000
```

Observed cancer rate was `r round(cancer_rate, 5)`.

Observed cancer incidence rate (CIR) was `r round(cir, 1)`

## Question 4

```{r}
expected <- 326/100000*person_years
expected
```

> The observed count of 67 is actually lower than the expected count, meaning 67 is not indicative of a cancer cluster in Randolph. 

## Question 5

```{r}
person_years2 <- 1138*12.5
expected2 <- 326/100000*person_years2
expected2
```

> 67 does now seem higher than the expected count of `r round(expected2, 1)`. We should incorporate statistical uncertainty before determining whether this is unusually high. 

## Question 6

> Assumption A could be false if family members within a household are systematically more or less likely than their neighbors to get screened, perhaps due to a famliy history of cancer. More screening means higher likelihood of discovering cancer. 

## Question 7

> Different ages might have different probability of diagnosis. Younger people are less likely to get cancer than older people.

## Question 8

> Different familial disease history and wealth might influence how likely people are to visit a hospital and receive a diagnosis.

## Question 9

```{r}
years <- 25
n <- 1138
person_years <- n*years
incidence_rate <- 326/100000

#simulate cases for a single neighborhood
rbinom(1, size = person_years, 
       prob = incidence_rate)

#simulate 10,000 neighborhoods
sims <- rbinom(10000, size = person_years,
               prob = incidence_rate)
#view summary of simulations
#sanity check- should get mean & median around 93
summary(sims)

#plot simulations
sims <- data.frame(cases = sims)
ggplot(sims, aes(x = cases)) +
  geom_histogram(color = "white")

#compute p-value
sum(sims$cases > 67)/nrow(sims)
```

## Question 10

```{r}
years <- 12.5
n <- 1138
person_years <- n*years
incidence_rate <- 326/100000

#simulate cases for a single neighborhood
rbinom(1, size = person_years, 
       prob = incidence_rate)

#simulate 10,000 neighborhoods
sims <- rbinom(10000, size = person_years,
               prob = incidence_rate)
#view summary of simulations
summary(sims)

#plot simulations
sims <- data.frame(cases = sims)
ggplot(sims, aes(x = cases)) +
  geom_histogram(color = "white")

#compute p-value
sum(sims$cases > 67)/nrow(sims)
```

## Skip 11 - 14

## Question 15

```{r}
glimpse(cancer)
bga_data <- cancer %>% 
  filter(location == "BGA")

m15 <- glm(cases ~ median_age + offset(log(person_years)), 
          data = bga_data, family = "poisson")
tidy(m15)
```

## Question 16

```{r}
b0 <- tidy(m15)[1,2] %>% pull()
b1 <- tidy(m15)[2,2] %>% pull()
rate_est <- exp(b0 + b1*62)
rate_est
cir_est <- rate_est*100000
cir_est

cir_obs <- bga_data %>% 
  filter(median_age == 62) %>% 
  pull(rate)
cir_obs

cir_obs - cir_est
```


> The estimated CIR is `r round(cir_est, 1)` and the observed CIT is `r round(cir_obs, 1)`. The estimate is not very good (the values are not close). 

## Question 17

```{r}
ctr_data <- cancer %>% 
  filter(location == "CTR")

m17 <- glm(cases ~ median_age + offset(log(person_years)), 
          data = ctr_data, family = "poisson")
tidy(m17)
```

> The $b_1$ estimate is larger for the CTR dataset, indicating that the log(cancer rate) is growing faster with age for this sample. 

## Question 18

```{r}
#extract regression coefficients for m17
b0_m17 <- tidy(m17)[1,2] %>% pull()
b1_m17 <- tidy(m17)[2,2] %>% pull()

#compute CIR for CTR data (m17), age = 62
cir_est_m17_62 <- exp(b0_m17 + b1_m17*62)*100000
cir_est_m17_62

#extract observed CIR for CTR data, age = 62
cir_obs_62 <- ctr_data %>% 
  filter(median_age == 62) %>% 
  pull(rate)
cir_obs_62
```

> For median age of 62 years:

> the estimated cancer incidence rate (CIR) is $\hat{\lambda/t} = exp(`r b0_m17` + `r b1_m17`(62))*100000 = `r cir_est_m17_62`$ 

> the observed CIR is `r cir_obs_62`

> these two values are fairly close (difference of `r cir_est_m17_62 - cir_obs_62`)

```{r}
#compute CIR for CTR data (m17), age = 72
cir_est_m17_72<- exp(b0_m17 + b1_m17*72)*100000
cir_est_m17_72

#extract observed CIR for CTR data, age = 72
cir_obs_72 <- ctr_data %>% 
  filter(median_age == 72) %>% 
  pull(rate)
cir_obs_72
```

> For median age of 62 years:

> the estimated cancer incidence rate (CIR) is $\hat{\lambda/t} = exp(`r b0_m17` + `r b1_m17`(72))*100000 = `r cir_est_m17_72`$ 

> the observed CIR is `r cir_obs_72`

> these two values are fairly close (difference of `r cir_est_m17_72 - cir_obs_72`)

## Question 19

```{r}
exp(b1_m17*10)
```

Cancer incidence rates are estimated to be `r round(exp(b1_m17*10),2)` times higher for each additional 10 years in median age. 

## Question 20

```{r}
m20 <- glm(cases ~ median_age + offset(log(person_years)), 
          data = cancer, family = "poisson")
tidy(m20)
```

> This model probably does not make sense because the estimated coefficients for median age are very different for the two locations. 

## Question 21

```{r}
m21 <- glm(cases ~ location + offset(log(person_years)), 
          data = cancer, family = "poisson")
tidy(m21)

b1_m21 <- tidy(m21)[2,2] %>% pull()
exp(b1_m21)
```

> The CIR is `r round(exp(b1_m21), 2)` times higher for CTR compared to BGA. 

## Question 22

```{r}
m22 <- glm(cases ~ median_age + location + 
             offset(log(person_years)), 
          data = cancer, family = "poisson")
tidy(m22)

b1_m22 <- tidy(m22)[2,2] %>% pull()
b2_m22 <- tidy(m22)[3,2] %>% pull()
exp(b1_m22)
exp(b2_m22)
```

> Adjusting for location, the CIR is estimated to increase by `r round(exp(b1_m22), 2)` times on average for each additional year in median age.

> Adjusting for median age, the CTR location is estimated to be `r round(exp(b2_m22), 2)` times higher than the BGA location. 

## Question 23

> Adjusting for median age, the CTR location is estimated to be `r round(exp(b2_m22), 2)` times higher than the BGA location. 

## Question 24

> The CTR data are not linear - they appear to follow an exponential distribution

## Question 25

```{r}
ggplot(cancer, aes(x = median_age,
                   y = log(rate),
                   group = location,
                   color = location,
                   linetype = location)) +
  geom_line() +
  geom_point()
```

> The CTR data are now linear but hte BGA data are not

## Question 26

> No, we cannot use the same general discription for both locations. 

## Question 27

```{r}
m27 <- glm(cases ~ median_age*location + 
             offset(log(person_years)), 
          data = cancer, family = "poisson")
tidy(m27)
```



## Question 28

```{r}
b0_m27 <- tidy(m27)[1,2] %>% pull()
b1_m27 <- tidy(m27)[2,2] %>% pull()
b2_m27 <- tidy(m27)[3,2] %>% pull()
b3_m27 <- tidy(m27)[4,2] %>% pull()
```

$$BGA \text{ (location = 0)}: \hat{\lambda}/t = exp(b_0 + b_1x) = exp(`r round(b0_m27,3)` + `r round(b1_m27,3)`x)$$

$$CTR \text{ (location = 1)}: \hat{\lambda}/t = exp((b_0 + b_2) + (b_1 + b_3)x) = exp(`r round(b0_m27 + b2_m27,3)` + `r round(b1_m27 + b3_m27,3)`x)$$

## Question 29

```{r}
exp((b1_m27 + b3_m27)*10) #ctr
exp((b1_m27)*10) #bga
```

> CTR location: exp(`r round(b1_m27 + b3_m27, 3)`*10)  = `r round(exp((b1_m27 + b3_m27)*10), 2)` times higher

> BGA location: exp(`r round(b1_m27, 3)`*10)  = `r round(exp((b1_m27)*10), 2)` times higher

## Question 30

```{r}
tidy(m27)
```

> yes, the relevant p-value is the one for the interaction term. Since it is small, there is evidence to suggest the location does affect the relationship between age and log-CIR. 

## Question 31

```{r}
summary_restricted <- summary(m22) #no interaction
summary_restricted
summary_full <- summary(m27) #interaction
summary_full
G <- summary_restricted$deviance - summary_full$deviance
G
df <- summary_restricted$df.residual - summary_full$df.residual
pchisq(G, df, lower.tail = FALSE)
```

> The LRT G-statistic is `r round(G,2)`, which follows a chi-squared distribution with `r df` degree of freedom. The p-value is `r round(pchisq(G, df, lower.tail = FALSE), 3)`, therefore there is sufficient evidence to conclude there is a significant interaction between age and location. The sample size is large enough here because all predicted means $\hat{\lambda}_i$ are larger than 5, as shown below.

```{r}
cancer <- cancer %>% 
  mutate(location_01 = if_else(location == "BGA", 0, 1),
         lambda_hat_m27 = person_years*exp(b0_m27 + b1_m27*median_age + b2_m27*location_01 + b3_m27*median_age*location_01))
cancer %>% pull(lambda_hat_m27)
```


## Question 32

> A quadratic model may be more appropriate, but with only 4 data points it's hard to discern. 

## Question 33

```{r}
ctr_data <- 
m33 <- glm(cases ~ median_age + I(median_age^2) + offset(log(person_years)), 
          data = ctr_data, family = "poisson")
summary(m33)
deviance_stat <- summary(m33)$deviance
df <- summary(m33)$df.residual
pchisq(deviance_stat, df, lower.tail = FALSE)
```

> With a p-value of `r round(pchisq(deviance_stat, df, lower.tail = FALSE), 3)`, there is not evidence that the model is a poor fit (see page 269). 

## Question 34

```{r}
summary(m27)
pchisq(summary(m27)$deviance, 
       df = summary(m27)$df.residual,
       lower.tail = FALSE)
```

> The deviance on the interaction model in 27 is `r round(summary(m27)$deviance, 1)` with `r summary(m27)$df.residual` degrees of freedom, yielding a p-value of `r pchisq(summary(m27)$deviance, df = summary(m27)$df.residual, lower.tail = FALSE)`. This indicates the model may not be a good fit. Looking at the individual deviance residuals shows that some of them are larger than 2 in magnitude. A plot indicates a quadratic term may be appropriate. 

```{r}
summary(m27)$deviance.resid

cancer <- cancer %>% 
  mutate(dev_res_m27 = summary(m27)$deviance.resid)
ggplot(cancer, aes(x = median_age, 
                   y = dev_res_m27)) +
  geom_point()
```

To conduct a LRT for whether a quadratic term is appropriate, we fit a new model with an interaction term and a quadratic term and compare it to `m27`, which is the interaction model without the quadratic term. 

```{r}
m34 <- glm(cases ~ median_age*location + I(median_age^2) +
             offset(log(person_years)), 
          data = cancer, family = "poisson")
summary_restricted <- summary(m27) #no quadratic
summary_restricted
summary_full <- summary(m34) #quadratic
summary_full
G <- summary_restricted$deviance - summary_full$deviance
G
df <- summary_restricted$df.residual - summary_full$df.residual
pchisq(G, df, lower.tail = FALSE)
```

> There is significant evidence that a quadratic term should be included in the model. 

# Galapagos

## Question 50

```{r}
gala <- read_csv("./data/C8 Gala.csv") %>% 
  clean_names()
glimpse(gala)

create_plot <- function(explanatory){
  ggplot(gala, aes(x = !!sym(explanatory),
                   y = log(species))) +
    geom_point()
}
create_plot("area")
create_plot("elevation")
create_plot("nearest")
create_plot("scruz")
create_plot("adjacent")
```

> Most of the covariates are not linearly related to the log(species) response variable, indicating a Poisson regression may not be best. 


## Question 51

```{r}
gala <- gala %>% 
  mutate(scruz = if_else(scruz == 0, scruz + 0.01, scruz))
create_plot2 <- function(explanatory){
  ggplot(gala, aes(x = log(!!sym(explanatory)),
                   y = log(species))) +
    geom_point()
}
create_plot2("area")
create_plot2("elevation")
create_plot2("nearest")
create_plot2("scruz")
create_plot2("adjacent")
m50 <- glm(cases ~ median_age*location + I(median_age^2) +
             offset(log(person_years)), 
          data = cancer, family = "poisson")
```

> Now the transfored explanatory variables do look linearly related to log(species), so a Poisson regression is more appropriate. log(area) and log(elevation) look like they will be the most useful predictors.

## Question 52

```{r}
m52 <- glm(species ~ log(area) + log(elevation), 
          data = gala, family = "poisson")
summary_m52 <- summary(m52)
summary_m52

G <- summary_m52$null.deviance - summary_m52$deviance
df <- summary_m52$df.null - summary_m52$df.residual
G
df
pchisq(G, df, lower.tail = FALSE)
```

> There is sufficient evidence to claim that at least one of log(area) and log(elevation) are useful predictors of log(species). 

## Question 53

```{r}
m53 <- glm(species ~ log(area) + log(elevation) + nearest + scruz + adjacent, 
          data = gala, family = "poisson")

summary_full <- summary(m53) #all 5 covariates
summary_restricted <- summary_m52 #2 covariates only

G <- summary_restricted$deviance - summary_full$deviance
df <- summary_restricted$df.residual - summary_full$df.residual
G
df
pchisq(G, df, lower.tail = FALSE)
```

> There is strong evidence that at least one of these additional covariates is useful. 

## Question 54

> Yes, log(elevation) appeared to be strongly related to log(species), but it is not significant in this updated model. 

## Question 55

```{r}
gala %>% 
  select(area, elevation, nearest, scruz, adjacent) %>% 
  cor()
```

> elevation and area are highly correlated

## Question 56

> Scruz and adjacent appear significant but did not appear related to log(species) in the graph. 

## Question 57

```{r}
summary_full
pchisq(summary_full$deviance, 
       summary_full$df.residual,
       lower.tail = FALSE)
```

> The deviance statistic is large and the p-value small, indicating that the model does not fit the data very well. 

## Question 58

```{r}
m58 <- glm(species ~ log(area) + log(elevation) + nearest + scruz + adjacent, 
          data = gala, family = "quasipoisson")
summary(m58)
```

> YOUR ANSWER HERE

## Question 59

```{r}
m58b <- glm(species ~ log(area) + log(elevation) + scruz + adjacent, 
          data = gala, family = "quasipoisson")
tidy(m58b)

m58c <- glm(species ~ log(area) + scruz + adjacent, 
          data = gala, family = "quasipoisson")
tidy(m58c)

m58d <- glm(species ~ log(area) + adjacent, 
          data = gala, family = "quasipoisson")
tidy(m58d)

b0 <- tidy(m58d)[1,2] %>% pull()
b1 <- tidy(m58d)[2,2] %>% pull()
b2 <- tidy(m58d)[3,2] %>% pull()

exp(b0 + b1*2 + b2*15)
```


