---
title: "AE-CH7"
author: "Your Name Here"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number_section: false
    highlight: "tango"
    theme: "cosmo"
    df-print: paged
editor: visual
editor_options: 
  chunk_output_type: console
---

This Application Exercise walks through the guided activities in Chapter 7 of the textbook. There are spaces for you to answer each question (with code and/or text). Feel free to also use this document to add your own notes along the way.

## Load packages & data

```{r}
#| label: load-packages-data
#| message: false
#| warning: false

library(tidyverse)
library(janitor)
library(ggridges)
library(broom)
library(cowplot)
#library(lme4)

shuttle <- read_csv("./data/C7 Shuttle.csv") %>% 
  clean_names()
glimpse(shuttle)
```

## Question 1

## Question 2

```{r}
ggplot(shuttle, aes(x = temperature,
                    y = success)) +
  geom_point()

ggplot(shuttle, aes(x = temperature,
                    y = factor(success))) +
  geom_boxplot()

ggplot(shuttle, aes(x = temperature,
                    y = factor(success))) +
  geom_density_ridges()
```

## Question 3

```{r}
ggplot(shuttle, aes(x = temperature,
                    y = success)) +
  geom_point() +
  geom_smooth(method = "lm")

m1 <- lm(success ~ temperature, 
         data = shuttle)
tidy(m1)

-1.90 + 0.0374*60
-1.90 + 0.0374*85
```

What's the problem with these predictions?

\[Skip Q4 - done in class together\]

## Question 5

```{r}


create_plot <- function(beta0, beta1){
  x <- seq(0, 35, 0.01)
  y <- exp(beta0 + beta1*x)/(1 + exp(beta0 + beta1*x))

  data <- data.frame(x, y)
  ggplot(data, aes(x = x, y = y)) +
    geom_line() +
    labs(title = paste("beta0 = ", beta0,
                       ", beta1 = ",
                       beta1, sep = ""))
}
p1 <- create_plot(-10, 0.5)
p2 <- create_plot(-5, 0.5)
p3 <- create_plot(-10, 1)
p4 <- create_plot(-5, 1)
p5 <- create_plot(-10, 1.5)
p6 <- create_plot(-5, 1.5)
plot_grid(p1, p3, p5, p2, p4, p6, nrow = 2)

p1 <- create_plot(10, -0.5)
p2 <- create_plot(5, -0.5)
p3 <- create_plot(10, -1)
p4 <- create_plot(5, -1)
p5 <- create_plot(10, -1.5)
p6 <- create_plot(5, -1.5)
plot_grid(p1, p3, p5, p2, p4, p6, nrow = 2)
```

## Question 6

```{r}
m2 <- glm(success ~ temperature, 
          family = "binomial",
          data = shuttle)
tidy(m2)
```

## Question 7

$$\hat{\pi}\_i = \frac{e^{b_0+b_1x_i}}{1 + e^{b_0+b_1x_i}}$$

```{r}
b0 <- tidy(m2)[1,2] %>% pull()
b1 <- tidy(m2)[2,2] %>% pull()
x <- c(31,50,75)
pi_hat <- exp(b0 + b1*x)/(1 + exp(b0 + b1*x))
pi_hat
```

## Question 8

```{r}
x <- c(60, 70)
pi_hat <- exp(b0 + b1*x)/(1 + exp(b0 + b1*x))
odds <- pi_hat/(1 - pi_hat)
odds
```

## Question 9

```{r}
exp(10*b1)
```

## Question 10

## Question 11

```{r}
b0_ols <- tidy(m1)[1,2] %>% pull()
b1_ols <- tidy(m1)[2,2] %>% pull()

b0_mle <- tidy(m2)[1,2] %>% pull()
b1_mle <- tidy(m2)[2,2] %>% pull()

data <- data.frame(x = seq(0,100, 1)) %>% 
  mutate(y_ols = exp(b0_ols + b1_ols*x)/(1 + exp(b0_ols + b1_ols*x)),
         y_mle = exp(b0_mle + b1_mle*x)/(1 + exp(b0_mle + b1_mle*x)))

ggplot(data = data) +
  geom_line(aes(x = x, y = y_ols), 
            color = "red") +
  geom_label(aes(x = 100, y = max(y_ols),
                 label = "OLS")) +
  geom_line(aes(x = x, y = y_mle), 
            color = "blue") +
  geom_label(aes(x = 100, y = max(y_mle),
                 label = "MLE")) +
  labs(x = "Temperature",
       y = "Probability of success",
       title = "MLE vs OLS estimation")


```

## Question 12

## Question 13

## Question 14

## Question 15

## Question 16

## Question 17

## Question 18
